\plaintitle

### 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
fleshing out 3 of the case studies

\begin{abstract}
  With growing attention on gig work
  --- ranging from the ``sharing economy'' to microtasks ---
  scholars have made connections to frameworks like Taylorism,
  and mechanisms such as worker advocacy and empowerment,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that our the underlying trend towards ``piecework''
  --- driven in part by the discretization,
  routenization,
  and external management of said work ---
  not only suggests, but in fact generates what we have observed:
  members of this transient workforce increasingly feeling
  disempowered,
  marginalized,
  and frustrated by the systems and platforms on which they work.

@@

\section{Introduction}
This section should outline the trend of contemporary gig work in the last 10 years,
in particular tracking the changes since
\citeauthor{crowdworkFuture} discussed
\citetitle{crowdworkFuture}
in \citeyear{crowdworkFuture}
\cite{crowdworkFuture}.
This would review the literature since then,
then point to the movements toward worker unionization, advocacy, etc\dots
that have happened in the last year or two.

In the reflection on the literature published in the last 5 years since
\citetitle{crowdworkFuture},
we notice a broader change in the form of work that's being done;
motivated by this realization,
we look at the emergent phenomena in piecework both as interconnected.
By looking at the movements toward
discretization,
routenization,
and management
as necessarily linked --- even causally --- to one another, 
we give ourselves a framing of contemporary piecework that explains
and arguably predicts
what we have seen thus far.

\subsection{Piecework as a lens to understand gig work}
It's not new that researchers have discussed gig work as an extension or resurgence
of piecework, but it is sufficiently important for the framing of this argument that
it merits a more direct treatment and some explanation here.
\citeauthor{crowdworkFuture}
investigate the future of crowd work by
situating and interrogating it through ``piecework'',
a term almost lost to history, but which for a time described
work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture}.
This work was largely in textiles and was predominantly done by women,
which by itself is a topic worthy of unpacking
\cite{scott1975women}.
Given the scope, however,
we will have to leave this complex topic with the superficial takeaway that
piecework and microtasks share some structural similarities:
\begin{inlinelist}
\item the work is done in the home;
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
\end{inlinelist}

@@

This paper will argue that the similarity is more than striking;
it was an inevitable milestone in the maturation of the type of work that is reemerging today.
\textit{How to continue???}

@@

\subsection{Context}
Importantly, since the reemergence of piecework we've seen substantial frustration
and resistance among the workers in this area
\cite{uberAlgorithm,turkopticon,dynamo}.
This paper attempts to make sense of the broader research on this piecework,
or ``gig work'', by framing this as one of several steps in the marginalization of workers,
starting with the discretization of tasks,
followed by routenization and the rise of workflows,
and finally the external management of workers.
All of this is to say that these milestones follow sequentially,
not coincidentally but necessarily,
and that by tracing this path using the corpus of scholarship on labor and workers
we can both make sense of past events and perhaps reasonably predict next steps.

We explore each topic
--- discretization, routenization, \& management ---
by looking at case studies in social computing.
Having validated this lens as a way of reasoning about contemporary piecework, or ``gig work'',
we turn to look ahead, envisioning future areas that researchers in social computing
--- and particularly digitally mediated work ---
should explore.

\section{Case Studies}
The past decade of research has shed light on crowd work from various perspectives,
and revealed 
\subsection{Discretization}
\textit{I want to argue that modern gig work is discretized,
and that this parallels some of the piecework that we're historically familiar with}
Work has become discretized in the last 10 years, particularly in the information work sector.
Breakthroughs like ``Find--fix--verify'', ``Cascade'', etc\dots
make it easier to assign smaller components of work to myriad workers and
recompose the constituent parts into something more complex
\cite{bernstein2015soylent,chilton2013cascade}.
Some work has even turned to ``break it down'',
inasmuch as decomposing what we might call ``macro'' tasks into microtasks
\cite{cheng2015break}
(\textit{there's a lot more to talk critique from this paper; sorry Justin.}).

@@

\subsection{Routenization}
The \textbf{discretization} and decomposition of work certainly facilitates
breaking tasks down,
parallelizing work,
and getting more done across a broader array of people;
but the key advantage in turning macro tasks 
into many micro tasks is that we can make that work \textit{routine}.

@@

Routenization can be seen in the contemporary in workflows
\cite{foundry,bernstein2015soylent}
[\textit{There are various papers that talk about creating a workflow;
maybe this is where FFV belongs? Foundry?}].

@@

\subsection{External Management}
\textbf{Discretization}, enabling \textbf{routenization}, thus allows us to
evaluate work in abstraction.
This means that we can

@@

\textit{The parallels here would be difficult to ignore;
Lily and others have pointed this out,
but it's worth thinking about the broader trend of
resisting management,
driven by routenization,
enabled by discretization.
Do that here.}

\subsection{Looking forward}
If we agree that
discretization,
routenization,
management,
and even resistance
necessarily follow one another according to this theoretical lens,
then we have to use it to attempt to envision
what comes next.

@@

\subsection{Routenization of work}
This leads into the next section, but I want to bring up the process of
making work about mass--manufacturing, at least inasmuch as
the instructions are the same for everyone.

@@

\subsection{Taylorism}
The routenization of work makes it possible to measure that process,
optimize it for certain characteristics, and ultimately
lead to Taylorism and scientific management.
This is not new;
researchers have studied and written about the slow creep of
algorithmic management and 
discretization \& routenization of work tasks.

>>>>

\begin{abstract}
  With growing attention on gig work
  --- ranging from the ``sharing economy'' to microtasks ---
  scholars have made connections to frameworks like Taylorism,
  and mechanisms such as worker advocacy and empowerment,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that our the underlying trend towards ``piecework''
  --- driven in part by the discretization,
  routinization,
  and external management of said work ---
  not only suggests, but in fact generates what we have observed:
  members of this transient workforce increasingly feeling
  disempowered,
  marginalized,
  and frustrated by the systems and platforms on which they work.

@@

\section{Introduction}

% \itshape
% What am I trying to say here?
% I want to get across that researchers have looked at gig work in various ways,
% but not come up with very persuasive overarching explanations for this whole process.
% That is, there's no broader theory making sense of why all of this is happening.

% Rather than make the argument that
% these things are all independently similar to the piecework
% (something to which researchers have already made the connection),
% I want to frame this as a huge process, involving:
% \begin{enumerate}
%   \item distribution of work,
%   \textbf{which leads to\dots}
%   \item discretization and routinization of work,
%   \textbf{which allows\dots}
%   \item management (scientific \& external),
%   \textbf{which causes workers to feel\dots}
%   \item alienation from work (Marx's idea) and resistance to management (also Marx probably but a more general labor advocacy theory too)
% \end{enumerate}
% okay, let's try this:
% \upshape

The past decade has seen
microtasks, the ``sharing economy'', and other instantiations of on--demand contract work
grow to occupy the fascination of both academic circles and our culture as a whole
\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
The research community has identified myriad connections between this emergent form of work
and the historically situated ``piecework'',
as well as a number of observations relating to
the frustration workers feel stemming from the management of this work
\cite{turkopticon,dynamo,uberAlgorithm}.
This body of work represents the cornerstones of how we understand crowd workers and on--demand labor more generally.

While much of this work appears to gesture toward
the parallels between contemporary on--demand work and piecework,
it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
This paper will attempt to do so by arguing that
the topics social computing researchers have investigated are
not just parallel to historical piecework and
the process of factorization that took place in the early \nth{20} century;
indeed, these phenomena are directly related, and that they precipitate
the social outcomes we have observed,
especially with regard to worker frustration, coordination, and resistance.

% This section should outline the trend of contemporary gig work in the last 10 years,
% in particular tracking the changes since
% \citeauthor{crowdworkFuture} discussed
% \citetitle{crowdworkFuture}
% in \citeyear{crowdworkFuture}
% \cite{crowdworkFuture},.
% In particular, we point to \citeauthor{crowdworkFuture}'s discussion of worthwhile
% questions to ask regarding \citetitle{crowdworkFuture}
% \cite{crowdworkFuture}.
% This would review the literature since then,
% then point to the movements toward worker unionization, advocacy, etc\dots
% that have happened in the last year or two.

In the reflection on the literature published in the last 5 years since
\citeauthor{crowdworkFuture}'s 
\citetitle{crowdworkFuture},
we notice a broader trend describing the change in work that's being done
% motivated by this realization,
% we look at the emergent phenomena in piecework both as interconnected
\cite{crowdworkFuture}.
Crowd--powered work is not just distributed across homes,
but dispersed among drivers, cleaners and other on--demand workers.
By looking at the movements toward
distributing work,
routinizing and breaking down tasks,
and externally managing workers
% discretization,
% routinization,
% and management
as linked to one another, 
we give ourselves a framing of contemporary piecework that explains,
and arguably predicts,
what we have seen thus far
--- and perhaps what we should expect to see going forward.

@@

This paper will argue that the familiarity is more than passing;
it was an inevitable milestone in the maturation of the type of work that is reemerging today.
``But'',
as \citeauthor{scholz2012digital} points out,
``it would be wrong to conclude that
in the realm of digital labor there is nothing new under the sun''
\cite{scholz2012digital}.
Indeed, we will attempt to reason about more than the larger--scale progression from
the discretization of work toward scientific management and future steps;
we will attempt to make sense of how the modern processes in digital labor,
in particular the use of the Internet as a medium to manage as well as \textit{do} work,
affect these unfolding events.

@@

\subsection{Piecework as a lens to understand gig work}
% It's not new to discuss gig work as an extension or even a resurgence of piecework
% \cite{crowdworkFuture,fixingChaos}.
% Nevertheless,
% it's sufficiently core to the argument here that a review is worthwhile.
\citeauthor{crowdworkFuture}
investigate the future of crowd work by
situating and interrogating it through ``piecework'',
a term almost lost to history, but which for a time described
work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture}.
This work was largely in textiles and was predominantly done by women,
which by itself is a topic worthy of unpacking
\cite{scott1975women}.
Given the scope, 
we'll leave this topic with the superficial takeaway that
piecework and microtasks share some structural similarities:
\begin{inlinelist}
\item the work is done in the home;
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
\end{inlinelist}
It may be said that
on--demand work shares, if not inspiration, then many of the same outcomes as piecework.

\subsection{Context}
Importantly, since the reemergence of piecework we've seen substantial frustration
and resistance among the workers in this area
\cite{uberAlgorithm,turkopticon,dynamo}.
This paper attempts to make sense of the broader research on this piecework,
or ``gig work'', by framing this as one of several steps in the marginalization of workers,
starting with the discretization of tasks,
followed by routinization and the rise of workflows,
and finally the external management of workers.
All of this is to say that these milestones follow sequentially,
not coincidentally but necessarily,
and that by tracing this path using the corpus of scholarship on labor and workers
we can both make sense of past events and perhaps reasonably predict next steps.

We explore each topic
--- discretization, routinization, \& management ---
by looking at case studies in social computing.
Having validated this lens as a way of reasoning about contemporary piecework, or ``gig work'',
we turn to look ahead, envisioning future areas that researchers in social computing
--- and particularly digitally mediated work ---
should explore.

\section{Case Studies}
The existing body of research has shed light on on--demand from various perspectives,
and revealed a number of topics that,
through our framing, are clearly situated together.
Those topics, at a very high level, are
\begin{enumerate}
\item the \textbf{distribution} of work,
\item the \textbf{routinization} and \textbf{discretization} of work, and finally
\item the \textbf{relationships} between the workers and the product of their work,
      as well as workers and managers.
\end{enumerate}

We will follow these topics, using them as prompts for case studies in
the emergence and
development of contemporary piecework.

\subsection{Distributing Work}
In both the historical and contemporary cases,
distributed, speculative work
% seemed a reasonable way
has promised
to diffuse the costs of capital--intensive resources
(whether they were sewing machines or laboratory equipment).

\textit{There's a reference to something that's \textbf{not} 99designs,
that describes a call for scientific solutions/designs to solve some problem
that the requesters have.
Honestly it's killing me that I can't remember the name,
and I need to find this before I do anything productive ever again.}

The HCI community is perhaps most familiar with examples such as
99designs,
Uber, Lyft,
and Amazon Mechanical Turk (AMT),
which all allow requesters in various forms to tap into
resources such as
cars,
computers, and above all
``cognitive surplus''
with relative ease
\cite{DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
This insight, that
workers can be geographically distributed,
% (especially given careful central management of those workers, discussed later),
has proven remarkably compelling
\cite{sensitiveTasks}.

The insight here seems to have been that
a substantial amount of work does not necessarily need to be done in collocated space,
but can be done from one's own home or other space arranged by the worker.
Moreover, by shifting the workplace into the worker's domain,
the cost of capital involved in that work can more easily be placed on workers' backs.
While yellow--cab organizations often manage drivers in similar ways to platforms such as Uber and Lyft
(inasmuch as drivers are often considered contractors, or otherwise exempt from the formal status of ``employee''),
these new systems go further,
by framing ownership and responsibility for costs
(e.g. automotive maintenance)
on drivers.
In our fieldwork, we discovered many longtime ``on--demand'' drivers
(affiliated with platforms such as Uber and Lyft)
who readily voiced frustration with the costs of
frequent oil changes and the unexpected replacement of expensive parts.

Distributing labor had similar effects on pieceworkers at the turn of the \nth{20} century;
textile workers were given the source material with which to work,
but were generally expected to use their own sewing materials and work in their own homes
\cite{hapke2004sweatshop}.

% the change seemed to begin with
% the redistribution of work into the homes and onto the diffuse capital of workers.
% 99designs may represent one of the first illustrative cases of
% speculative, distributed labor in the contemporary form,
% but speculative labor and even on--demand design work is not necessarily new;
% \cite{99designs}.

% distributed; uber, etc... ->
% routinization; crowdwork - > optimization
% - > accommodation and resistance, alienation?
% tom malone

% This framing of labor worked for some time in the late \nth{19} and early \nth{20} centuries,

% This work is only made to seem new by framing the units of work as the output of ``task--completion APIs'',
% making it possible
% (and indeed advantageous)
% to allocate work on the fly to people in different places
% \cite{storiesIraniSilberman}.

\subsection{Discretization and Routinization}

% I never talk about factorization it's too soon ahhhh
With piecework and factorization making mass manufacturing more viable,
businesses could produce at scales that were previously unimaginable.
Nevertheless, this approach to labor and the production capacity it offered was not without its hurdles.
One of the issues that challenged mass production at the turn of the \nth{20} century,
as well as today in crowd work and elsewhere,
has been that of ensuring that the quality of the work is sufficiently good.
One of the

% hey so i need to bring in citations on piecework for this stuff because I didn't exactly live through this time

% is ANY of this good?

% The approach to piecework described above hinged in some sense on
% the shared understanding of the work to be done;
% workers generally understood what was required of sewing together pieces of fabric.
% Still, the work was speculative in that inadequately sewn clothing would not earn payment.
% This represented a risk for pieceworkers,
% who might spend time working on 10 garments only to be paid for 2 or 3.
% It also represented a potential cost for management,
% who would need to salvage the material
% (if it was even possible),
% and give that material to another pieceworker to do it correctly.

% i need to find the source for this, because it's honestly only a vague quote in my head,
% so... maybe i imagined it?

% One of the solutions that

% Broadly speaking, the solution to this problem is to make instructions clearer and easier to follow.
% Further, by establishing a shared understanding
% --- a ``collective conscious'' or even a corporate culture regarding work ---

\textit{I want to argue that modern gig work is discretized,
and that this parallels some of the piecework that we're historically familiar with}

Work has become discretized in the last 10 years, particularly in the information work sector.
Advances such as
``find--fix--verify'' % (described in \citetitle{bernstein2015soylent})
and
``Cascade'' % (described in \citetitle{chilton2013cascade})
make it easier to assign smaller components of work to myriad workers and
recompose the constituent parts into something more complex
\cite{bernstein2015soylent,chilton2013cascade}.
Some work has taken special interest in the process of breaking tasks down into more manageable parts
% Some work has even turned to break it down'',
inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
both for the purposes of crowd--sourcing in the sense of out--sourcing work
as well as ``selfsourcing''
\cite{cheng2015break,teevan2014selfsourcing}.

@@

% \subsection{routinization}
The \textbf{discretization} and decomposition of work certainly facilitates
breaking tasks down,
parallelizing work,
and getting more done across a broader array of people;
but the key advantage in turning macro tasks 
into many micro tasks is that we can make that work \textit{routine}.

@@

routinization can be seen in the contemporary in workflows
\cite{foundry,bernstein2015soylent}
[\textit{There are various papers that talk about creating a workflow;
maybe this is where FFV belongs? Foundry?}].

@@

\subsection{External Management}
\textbf{Discretization}, enabling \textbf{routinization}, thus allows us to
manage work more meaningfully.
Routinization in particular made it possible
(and indeed preferable) to assume that people
% as \citeauthor{hu1961parallel} describes the premise of this sort of work,
``\dots are of equal ability and \dots can do any of the $n$ jobs''
\cite{hu1961parallel}.
\citeauthor{storiesIraniSilberman}
find similarly that
this work is only made to seem new by framing the units of work as the output of ``task--completion APIs''
% making it possible
% (and indeed advantageous)
% to allocate work on the fly to people in different places
\cite{storiesIraniSilberman}.

Crucially,
\citeauthor{hu1961parallel}'s case study assumed that workers were similarly trained and capable,
which in \citeyear{hu1961parallel}
meant that workers would at least have to be trained in person, if not instructed as well.
Throughout the industrial revolution, workers flocked to centralized locales of production,
and physically situated factories made it both
difficult to ignore the general pattern of mistreatment and
relatively easy to organize.

% construction and manufacturing in various forms
% (from textiles to automobiles)
% encourage the consolidation of capital--intensive resources like factories.
% The prohibitively expensive nature of factory equipment may have been a key factor in 
\citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case

% evaluate work in abstraction.
% This means that we can

@@

\textit{The parallels here would be difficult to ignore;
Lily and others have pointed this out,
but it's worth thinking about the broader trend of
resisting management,
driven by routinization,
enabled by discretization.
Do that here.}

\subsection{Looking forward}
If we agree that
discretization,
routinization,
management,
and even resistance
necessarily follow one another according to this theoretical lens,
then we have to use it to attempt to envision
what comes next.

@@

\subsection{routinization of work}
This leads into the next section, but I want to bring up the process of
making work about mass--manufacturing, at least inasmuch as
the instructions are the same for everyone.

@@

\subsection{Taylorism}
The routinization of work makes it possible to measure that process,
optimize it for certain characteristics, and ultimately
lead to Taylorism and scientific management.
This is not new;
researchers have studied and written about the slow creep of
algorithmic management and 
discretization \& routinization of work tasks.
